Policy optimization for dynamic spatiotemporal systems \hspace{1em} \insertframenumber

### 2014-11-10 pot/LaTeX/skewnormal.tex
tweaks for time series

The paper is organized as follows.
In section \ref{s:spatial}, we describe a spatial skew-$t$ process with partitioning and thresholding.
In section \ref{s:temporal}, we provide a way to model temporal correlation.
The computing is described in section \ref{s:comp}.
In section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared with a na{\"i}ve Gaussian method.
We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from the eastern US in section \ref{s:analysis}.
The final section provides brief discussion and direction for future research.

@@

\subsection{Extremal dependence}
One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
The $\chi$ statistic for the upper tail is given by
\begin{align*}
  \chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c).
\end{align*}
In a stationary spatial process, we can write the $\chi$ coefficient as
\begin{align*}
  \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(0) > c | Y(h) > c)
\end{align*}
where $h = ||\bs_1 - \bs_2||$.
If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
For Gaussian processes, $\chi(h) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
However, for the skew-$t$ process described in \ref{s:data}, $\chi(h) > 0$ \citep[see Appendix A.5 for the expression]{Padoan2011}.

@@

Consider a set of spatial knots $\bw_{k} \sim \text{Uniform}(\calD)$ that define a random daily partition $P_{1}, \ldots, P_{K}$ of the spatial domain of interest $\calD \subset \calR^2$ such that
\begin{align*}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align*}
Then, the observation at each site becomes
\begin{align}
  Y(\bs) &= X(\bs) \bbeta + \alpha |z(\bs)| + \sigma v(\bs),
\end{align}
and, for $\bs \in P_{k}$,
\begin{align}
  z(\bs) &= z_{tk} \label{eq:sitez}\\
  \sigma^2(\bs) &= \sigma^2_{k} \label{eq:sitesig}.
\end{align}
So, within each partition, $\bY$ follows the distribution given in (\ref{eq:fullmodel}).

When incorporating the random daily partition, the $\chi$ statistic becomes
\begin{align}
  \chi(h) = \pi(h) \chi(h) + (1 - \pi(h)) \chi(h)
\end{align}
where $\pi(h)$ is the probability that two sites separated by distance $h$ are in the same partition.
Asymptotic dependence is eliminated as $h$ increases, because
\begin{align}
  \lim_{h \rightarrow \infty} \chi(h) = \lim_{h \rightarrow \infty} \pi(h) \chi(h) = 0.
\end{align}
A proof of this can be found in Appendix A.3.

@@

\subsection{Accounting for temporal dependence} \label{s:temporal}
In a standard block-maxima approach, the temporal dependence is assumed to be negligible because blocks (e.g. yearly maxima) are taken to be large enough to assume independence.
However, when using daily meausrements, the assumption of temporal independence is no longer appropriate.
We account for temporal dependence through an AR(1) time series on the spatial knots, $z_t(\bs)$ terms, and $\sigma^2_t(\bs)$ terms.
We first, transform the spatial knots from $\calD$ to $\calR^2$ as follows.
Let
\begin{align*}
  \bw^*_k = \Phi^{-1}_2\left[ \frac{ (\bw_k - \min(\bs))}{ \text{range}(\bs) } \right].
\end{align*}
where $\Phi_2$ represents a bivariate standard normal density function.
Then $\bw^*_k \in \calR^2$.
We use a copula on the $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
Let
\begin{align*}
  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
\end{align*}
where $\Phi$ is a univariate standard normal density function, and IG is the distribution function for an IG$(a, b)$ random variable.
Then, when $t > 1$ the time series is modeled as
\begin{align}
  \bw^*_{t, k} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
  z_{t, k} | z_{t-1, k} &\sim N \left[\phi_z z_{t-1, k}, \sigma^2_{t,k} (1 - \phi_z^2)\right] \\
  \sigma^{2*}_{t, k} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
\end{align}
where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
All three of the time series are initialized using $N(0, 1)$.

\subsection{Hierarchical model}\label{s:hier}
Conditioned on $z_{t,k}(\bs) \iid $ $N(0, \sigma^2_{t,k})$, $\sigma^2_{t,k}(\bs) \sim IG(a, b)$, and $P_{t,k}$, the marginal distributions are skew-$t$ and the joint distribution within a partition is multivariate skew-$t$.
However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
We model this with a Bayesian hierarchical model as follows.
Let $\bw_{t,1}, \ldots, \bw_{t,K}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, so that
\begin{align*}
 P_{t,k} = \{\bs : k = \argmin_\ell || \bs_t - \bw_{t\ell} ||\}.
\end{align*}

Then
\begin{align}
   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{t,k}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
   z_t(\bs) &= z_{t,k} \text{ if } \bs \in P_{t,k}\\
   \sigma^2_{t}(\bs) &= \sigma^2_{t,k} \text{ if } \bs \in P_{t,k}\\
   \alpha &\sim N(0, 10)\\
   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
   z_{t,k} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{t,k} (1 - \phi_z^2))\\
   \sigma^{2*}_{t,k} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
   \bw^*_{t,k} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
\end{align}
where $\Theta = \{\rho, \nu, \gamma\}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$, and the proportion of variance accounted for by the spatial variation is $\gamma$.

\section{Computation}\label{s:comp}
The MCMC for this model is fairly straightforward.
First, we impute values below the threshold.
Then, we update $\Theta$ using random walk MH or Gibbs sampling when appropriate.
Finally, we make spatial predictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z_{tl}$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.

We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
After conditioning on $\alpha$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \Sigma)$.
After imputing the censored observations, we update the model parameters.
To update the model parameters, we use standard Gibbs updates for parameters when possible.
In the case Gibbs sampling is not possible, parameters are updated using a random-walk Metropolis Hastings algorithm.
See Appendices A.1 and A.2 for details regarding the MCMC.
The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $(\bs^*)$.
This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.

@@

\subsection{Design}\label{s:simdesign}
For all simulation designs, we generate data from the model presented in Section \ref{s:part} using $n_s=130$ sites and $n_t=50$ independent days.
The sites are generated Uniform$([0, 10] \times [0, 10])$.
We generate data from 6 different simulation designs:
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knot
  \item $t$ marginal, $K=1$ knot
  \item $t$ marginal, $K=5$ knots
  \item skew-$t$ marginal, $K=1$ knots
  \item skew-$t$ marginal, $K=5$ knots
  \item Max-stable.
\end{enumerate}
In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter, $\nu = 0.5$, and spatial range, $\rho = 0.1$.
For the covariance matrices in designs 1 -- 5, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
In the first design, $\sigma^2 = 2$ is used for all days.
For designs 2 -- 4, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
For designs 1 -- 3, we set $\alpha = 0$.
For designs four and five, $\alpha = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
In all six designs, the mean ($\mu(\bs)$) is assumed to be constant across space.

$M = 50$ data sets are generated for each design.
For each data set we fit the data using
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knots
  \item skew-$t$ marginal, $K=1$ knots, $T=-\infty$
  \item $t$ marginal, $K=1$ knots, $T=q(0.80)$
  \item skew-$t$ marginal, $K=5$ knots, $T=-\infty$
  \item $t$ marginal, $K=5$ knots, $T=q(0.80)$
\end{enumerate}
where $q(0.80)$ is the 80th sample quantile of the data.
The design matrix $\bX$ includes an the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{U}(0, 1)$, $\log(\rho) \sim \text{N}(-2, 1)$.
The skewness parameter has prior $\alpha \sim \text{N}(0, 2)$.
The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.

\subsection{Cross validation}\label{s:modelselect}
Models were compared using cross validation with 100 sites used as training sites and 30 sites witheld for testing.
The model was fit using the training set, and predictions were generated at the testing site locations.
Because one of the primary goals of this model is to predict extreme events, we quantify use Brier scores to select the model that best fits the data \citep{Gneiting2007}.
The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
For the Brier score, a lower score indicates a better fit.

\subsection{Results}\label{s:simresults}
We compared the Brier scores for exceeding 11 different thresholds for each dataset.
For each threshold, Brier score were calculated for each day and site in the testing set.
The Brier scores were then averaged over all the days and sites to obtain a single Brier score for threshold $c$.
The thresholds used for the Brier scores are extreme quantiles from the simulated data.
Figure \ref{fig:simbrierscores} gives the Brier score relative to the Brier score for the Gaussian method calculated as
\begin{align*}
  \text{BS}_{\text{rel}} = \frac{\text{BS}_{\text{method}}}{\text{BS}_{\text{Gaussian}}}.
\end{align*}
For data that are relatively symmetric, the standard Gaussian method outperforms the other methods.
However, when the data come from a skewed distribution (settings 3 -- 6), the partitioned and thresholded methods outperform the Gaussian method.
Additionally, for both of the skew-$t$ settings ($K = 1$, $K = 5$), the skew-$t$ method with the appropriate number of partitions performs the best.
Finally, the simulation results suggest that for data from a skew-$t$ distribution, thresholded methods provide some improvement over non-thresholded models for the extreme thresholds.
With the exception of the Brier score for $c = q(0.995)$ in setting 6, the improvement over the Gaussian method is statistically significant using a Wilcoxon signed rank test ($\alpha = 0.05$).

@@

\section{Data analysis}\label{s:analysis}
To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 2005 at 735 Air Quality System (AQS) monitoring sites in the eastern United States as the response.
For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
Initially, we fit a linear regression assuming that the mean function can be expressed as
\begin{align}
  \mu_t(\bs) = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
\end{align}
The data from 10 July are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 8 d.f. and $\alpha = 1$.
\begin{center}
\begin{figure}
  \includegraphics[width=0.5\linewidth]{plots/ozone-10jul.pdf}
  \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
  \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals (right)}
  \label{fig:ozone}
\end{figure}
\end{center}
The $\chi(h)$ and $\chi(t)$ plots indicate that the residuals still demonstrate spatial and temporal dependence (see Figure \ref{fig:chi-st})
\begin{center}
\begin{figure}
  \includegraphics[width=0.5\linewidth]{plots/chi-h-ozone.pdf}
  \includegraphics[width=0.5\linewidth]{plots/chi-t-ozone.pdf}
  \caption{$\chi(h)$ plot for the residuals. The vertical line indicates distance after which observations no longer demonstrate dependence (left) $\chi(t)$ plot for the residuals (right)}
  \label{fig:chi-st}
\end{figure}
\end{center}

@@

\subsection*{Variance terms}
When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis Hastings update.
First, we generate from one of two candidate distributions.
In the case of temporal dependence, we generate a candidate for ${\sigma^*_{tk}}^2$ using a random walk Gaussian candidate distribution
\begin{align*}
  {\sigma^{*2}_{tk}}^{(c)} \sim \text{N}({\sigma^{*2}_{tk}}^{(r - 1)}, s^2)
\end{align*}
where ${{\sigma^*_{tk}}^2}^{(r - 1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
In the case where there is no temporal dependence, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution where $a^* = n_{tk} / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk}] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
The acceptance ratio is
\begin{align*}
  R = \left\{
    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
    \right\} \times \left\{
    \frac{ l[z_tk | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[z_tk | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
    \right\} \times \left\{
    \frac{ p[ {\sigma^2_{tk}}^{(c)} ] }
         { p[ {\sigma^2_{tk}}^{(r - 1)}] }
    \right\} \times \left\{
    \frac{ c[ {\sigma^2_{tk}}^{(r - 1)}] }
         { c[ {\sigma^2_{tk}}^{(c)}]}
    \right\}
\end{align*}
where $p[\cdot]$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
The candidate is accepted with probability $\min\{R, 1\}$.

>>>>

The paper is organized as follows.
In section \ref{s:spatial}, we describe a spatial skew-$t$ process with partitioning and thresholding.
In section \ref{s:temporal}, we expand the model to incorporate temporal dependence.
The computing is described in section \ref{s:comp}.
In section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared with a na{\"i}ve Gaussian method.
We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from the eastern US in section \ref{s:analysis}.
The final section provides brief discussion and direction for future research.

@@

\subsection{Extremal dependence}
One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
The $\chi$ statistic for the upper tail is given by
\begin{align*}
  \chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c).
\end{align*}
In a stationary spatial process, we can write the $\chi$ coefficient as
\begin{align*}
  \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(0) > c | Y(h) > c)
\end{align*}
where $h = ||\bs_1 - \bs_2||$.
If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
For Gaussian processes, $\chi(h) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
However, for the skew-$t$ process described in \ref{s:data}, $\chi(h) > 0$ \citep[they give an expression that is the sum of two survival functions from the skew-$t$ distribution. The expression isn't terribly informative, but I can include it if you think it would be helpful.]{Padoan2011}.

@@

Consider a set of spatial knots $\bw_{k} \sim \text{Uniform}(\calD)$ that define a random daily partition $P_{1}, \ldots, P_{K}$ of the spatial domain of interest $\calD \subset \calR^2$ such that
\begin{align*}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align*}
Then, the observation at each site becomes
\begin{align}
  Y(\bs) &= X(\bs) \bbeta + \alpha |z(\bs)| + \sigma v(\bs),
\end{align}
and, for $\bs \in P_{k}$,
\begin{align}
  z(\bs) &= z_{tk} \label{eq:sitez}\\
  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig}.
\end{align}
So, within each partition, $\bY$ follows the distribution given in (\ref{eq:fullmodel}).

When incorporating the random daily partition, for two sites in the same partition, the $\chi$ statistic is $\chi(h) = \chi_{\text{skew-}t}(h)$. However, for two sites in different partitions, the $\chi$ statistic should be the same as the Gaussian $\chi(h)$, because the only correlation between the two sites, is the correlation given by $\bSigma$.
So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$
where $\pi(h)$ is the probability that two sites separated by distance $h$ are in the same partition.
A proof of this is given in Appendix A.3.

@@

\section{Accounting for temporal dependence} \label{s:temporal}
In a standard block-maxima approach, the temporal dependence is assumed to be negligible because blocks (e.g. yearly maxima) are taken to be large enough to assume independence.
However, when using daily meausrements, the assumption of temporal independence is no longer appropriate.
There are several places where temporal dependence could be included, including the residual $v(\bs)$.
However, we choose to allow for temporal dependence in the $\bw$, $z(\bs)$, and $\sigma(\bs)$ terms because these terms derive the tail behavior.
Specifically, we incorporate an AR(1) time series.
We first, transform the spatial knots from $\calD$ to $\calR^2$ as follows.
Let
\begin{align*}
  \bw^*_k = \Phi^{-1}_2\left[ \frac{ (\bw_k - \min(\bs))}{ \text{range}(\bs) } \right].
\end{align*}
\hl{What's the best way to write the $\bw$ transformation as 1-D? $w_x$ and $w_y$ or $w_1$ and $w_2$?}
where $\Phi_2$ represents a bivariate standard normal density function.
Then $\bw^*_k \in \calR^2$.
We use a copula on the $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
Let
\begin{align*}
  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
\end{align*}
where $\Phi$ is a univariate standard normal density function, and IG is the distribution function for an IG$(a, b)$ random variable.
Then, when $t > 1$ the time series is modeled as
\begin{align*}
  \bw^*_{1, k} & \sim N_2(0, 1)\\
  z_{1k} & \sim N(0, \sigma^2_{1k})\\
  \sigma^{2*}_{1k} &\sim N(0, 1)\\
  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
  z_{tk} | z_{t-1, k} &\sim N \left[\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
\end{align*}
where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
These are stationary time series models with marginal distributions
\begin{align*}
  \bw^*_{k} & \sim N_2(0, 1)\\
  z_{k} & \sim N(0, \sigma^2_{k})\\
  \sigma^{2*}_{k} &\sim N(0, 1).
\end{align*}
For each day, the model is identical to the spatial-only model in Seciton \ref{s:part}.

\subsection{Hierarchical model}\label{s:hier}
Conditioned on $z_{tk}(\bs)$, $\sigma^2_{t,k}(\bs)$, and $P_{t,k}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
We model this with a Bayesian hierarchical model as follows.
Let $\bw_{t,1}, \ldots, \bw_{t,K}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in \ref{s:part}.

Then
\begin{align}
   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
   \alpha &\sim N(0, 10)\\
   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
\end{align}
where $\Theta = \{\rho, \nu, \gamma\}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$, and the proportion of variance accounted for by the spatial variation is $\gamma$.

\section{Computation}\label{s:comp}
First, we impute values below the threshold.
Then, we update $\Theta$ using Metropolis Hastings or Gibbs sampling when appropriate.
Finally, we make spatial predictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z(\bs)$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.

We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
After conditioning on $\alpha$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \bSigma)$.
After imputing the censored observations, we update the model parameters.
To update the model parameters, we use standard Gibbs updates for parameters when possible.
In the case Gibbs sampling is not possible, parameters are updated using a random-walk Metropolis Hastings algorithm.
See Appendices A.1 and A.2 for details regarding the MCMC.
The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.

@@

\subsection{Design}\label{s:simdesign}
For all simulation designs, we generate data from the model in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
The sites are generated Uniform$([0, 10] \times [0, 10])$.
We generate data from 6 different simulation designs:
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knot
  \item $T$ marginal, $K=1$ knot
  \item $T$ marginal, $K=5$ knots
  \item Skew-$t$ marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=5$ knots
  \item Max-stable.
\end{enumerate}
In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 0.1$.
For the covariance matrices in designs 1 -- 5, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
In the first design, $\sigma^2 = 2$ is used for all days.
For designs 2 -- 4, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
For designs 1 -- 3, we set $\alpha = 0$.
For designs four and five, $\alpha = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
\hl{I know I need more here about the max-stable method}
In all six designs, the mean $\mu(\bs) = 10$ is assumed to be constant across space.

$M = 50$ data sets are generated for each design.
For each data set we fit the data using
\begin{enumerate} \setlength{\itemsep}{-0.5em}
  \item Gaussian marginal, $K=1$ knots
  \item Skew-$t$ marginal, $K=1$ knots, $T=-\infty$
  \item $T$ marginal, $K=1$ knots, $T=q(0.80)$
  \item Skew-$t$ marginal, $K=5$ knots, $T=-\infty$
  \item $T$ marginal, $K=5$ knots, $T=q(0.80)$
\end{enumerate}
where $q(0.80)$ is the 80th sample quantile of the data.
The design matrix $\bX$ includes an the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{U}(0, 1)$, $\log(\rho) \sim \text{N}(-2, 1)$.
The skewness parameter has prior $\alpha \sim \text{N}(0, 2)$.
The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.
The knots have priors $\bw \sim \text{Unif}(\calD)$.
We do not fit the data using the max-stable methods from \citet{Reich2012} because of the time it takes.

\subsection{Cross validation}\label{s:modelselect}
Models were compared using cross validation with 100 sites used as training sites and 30 sites witheld for testing.
The model was fit using the training set, and predictions were generated at the testing site locations.
Because one of the primary goals of this model is to predict extreme events, we use Brier scores to select the model that best fits the data \citep{Gneiting2007}.
The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
We average the Brier scores over all test sites and days.
For the Brier score, a lower score indicates a better fit.

\subsection{Results}\label{s:simresults}
We compared the Brier scores for exceeding 11 different thresholds for each dataset.
The thresholds used for the Brier scores are extreme quantiles from the simulated data.
Figure \ref{fig:simbrierscores} gives the Brier score relative to the Brier score for the Gaussian method calculated as
\begin{align*}
  \text{BS}_{\text{rel}} = \frac{\text{BS}_{\text{method}}}{\text{BS}_{\text{Gaussian}}}.
\end{align*}
For data that are relatively symmetric, the standard Gaussian method outperforms the other methods.
However, when the data come from a skewed distribution (settings 3 -- 6), the partitioned and thresholded methods outperform the Gaussian method.
Additionally, for both of the skew-$t$ settings ($K = 1$, $K = 5$), the skew-$t$ method with the appropriate number of partitions performs the best.
Finally, the simulation results suggest that for data from a skew-$t$ distribution, thresholded methods provide some improvement over non-thresholded models for the extreme thresholds.
With the exception of the Brier score for $c = q(0.995)$ in setting 6, the improvement over the Gaussian method is statistically significant using a Wilcoxon signed rank test ($\alpha = 0.05$).

@@

\section{Data analysis}\label{s:analysis}
To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 2005 at 735 Air Quality System (AQS) monitoring sites in the eastern United States as the response (see Figure \ref{fig:ozone}).
For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
Initially, we fit a linear regression assuming a mean function of
\begin{align}
  \mu_t(\bs) = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
\end{align}
The data from July 10 are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 8 d.f. and $\alpha = 1$.
\begin{center}
\begin{figure}
  \includegraphics[width=0.5\linewidth]{plots/ozone-10jul.pdf}
  \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
  \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals (right)}
  \label{fig:ozone}
\end{figure}
\end{center}
The $\widehat{\chi}(h)$ and $\widehat{\chi}(t)$ plots estimate the spatial and temporal dependence using sample data.
For $\widehat{\chi}(h)$, we take the sample proportion of all pairs of sites within distance $h$ of one another that are above a threshold $T$.
\hl{I know I need another sentence or two here about $\widehat{\chi}(t)$}
indicate that the residuals still demonstrate spatial and temporal dependence (see Figure \ref{fig:chi-st})
\begin{center}
\begin{figure}
  \includegraphics[width=0.5\linewidth]{plots/chi-h-ozone.pdf}
  \includegraphics[width=0.5\linewidth]{plots/chi-t-ozone.pdf}
  \caption{$\widehat{\chi}(h)$ plot for the residuals. The vertical line indicates the distance after which observations no longer demonstrate dependence (left) $\widehat{\chi}(t)$ plot for the residuals (right)}
  \label{fig:chi-st}
\end{figure}
\end{center}

@@

\subsection*{Variance terms}
When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis Hastings update.
First, we generate from one of two candidate distributions.
In the case of temporal dependence, we generate a candidate for ${\sigma^*_{tk}}^2$ using a random walk Gaussian candidate distribution
\begin{align*}
  {\sigma^{*2}_{tk}}^{(c)} \sim \text{N}({\sigma^{*2}_{tk}}^{(r - 1)}, s^2)
\end{align*}
where ${{\sigma^*_{tk}}^2}^{(r - 1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
In the case where there is no temporal dependence, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
The acceptance ratio is
\begin{align*}
  R = \left\{
    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
    \right\} \times \left\{
    \frac{ l[z_tk | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[z_tk | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
    \right\} \times \left\{
    \frac{ p[ {\sigma^2_{tk}}^{(c)} ] }
         { p[ {\sigma^2_{tk}}^{(r - 1)}] }
    \right\} \times \left\{
    \frac{ c[ {\sigma^2_{tk}}^{(r - 1)}] }
         { c[ {\sigma^2_{tk}}^{(c)}]}
    \right\}
\end{align*}
where $p[\cdot]$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
The candidate is accepted with probability $\min\{R, 1\}$.
